---
title: "Your Paper Title Here"
author: "Your Name(s)"
date: Date
output: pdf_document
urlcolor: blue
header-includes:
- \usepackage{amsmath,amsfonts,amssymb}
- \usepackage{setspace} \doublespacing
fontsize: 11pt
---

```{r setup, include=FALSE}
#Use this code chunk to include libraries, and set global options.
```

# Abstract
Write your abstract here.

# Keywords
*Keyword 1*, *Keyword 2*, *Keyword 3*, *Keyword 4*

\newpage

# Introduction / Background
Provide an introduction with background information.

```{r label="important_R_code", eval=TRUE, echo=FALSE}
# You can reference your code in the appendix (sample here).
```

# Statistical Methods

## Model
Describe the statistical model used.

## Likelihood Inference
Detail the likelihood approach.

## Likelihood Inference

Let $T$ denote the number of infections observed in the vaccine group among the
total $n=170$ infections recorded in the clinical trial. We assume

\[
T \sim \text{Binomial}(n,\pi),
\]

where $\pi = P(\text{infection occurs in the vaccine group} \mid \text{infection})$.
In the observed data, $t=8$ infections occurred in the vaccine group and
$162$ occurred in the placebo group.

The parameter of scientific interest is the vaccine efficacy $\psi$, defined as

\[
\psi = \frac{\pi_p - \pi_v}{\pi_p},
\]

where $\pi_v$ and $\pi_p$ denote the probabilities of infection in the vaccine
and placebo groups respectively. Under the binomial formulation of the trial
data, this relationship can be written in terms of $\pi$ as

\[
\psi = \frac{1-2\pi}{1-\pi}.
\]

---

### Likelihood Function

The likelihood function for $\pi$ is

\[
L(\pi)
= P(T=t|\pi)
= \binom{n}{t}\pi^t(1-\pi)^{n-t}.
\]

Taking the logarithm gives the log-likelihood function

\[
\ell(\pi)
= \log L(\pi)
= t\log(\pi) + (n-t)\log(1-\pi).
\]

The maximum likelihood estimator (MLE) is obtained by maximizing the likelihood
function. For the binomial model, the MLE has the closed form

\[
\hat{\pi}_{MLE} = \frac{T}{n}.
\]

Because maximum likelihood estimators are invariant under transformation,
the MLE for vaccine efficacy is

\[
\hat{\psi}_{MLE}
=
\frac{1-2\hat{\pi}}{1-\hat{\pi}}.
\]

---

### Large Sample Inference

Under regularity conditions and for sufficiently large sample sizes,
the MLE is approximately normally distributed:

\[
\hat{\pi}_{MLE}
\approx
N\left(\pi, \frac{\pi(1-\pi)}{n}\right).
\]

This result allows us to construct an approximate $(1-\alpha)$ confidence
interval for $\pi$ using

\[
\hat{\pi} \pm z_{\alpha/2}
\sqrt{\frac{\hat{\pi}(1-\hat{\pi})}{n}}.
\]

The endpoints of this interval can then be transformed using the relationship
between $\psi$ and $\pi$ to obtain a confidence interval for vaccine efficacy.

---

### Likelihood Ratio Test

To assess whether the vaccine efficacy meets the regulatory benchmark,
we test the hypothesis

\[
H_0 : \psi = 0.3
\qquad
\text{versus}
\qquad
H_1 : \psi \ne 0.3.
\]

The likelihood ratio test statistic is

\[
W = 2\left[\ell(\hat{\pi}) - \ell(\pi_0)\right],
\]

where $\pi_0$ corresponds to the null value of $\psi$ and is obtained from

\[
\pi_0 = \frac{1-\psi_0}{2-\psi_0}.
\]

Under the null hypothesis and standard regularity conditions,
the test statistic is approximately distributed as

\[
W \sim \chi^2_1.
\]

The corresponding $p$-value is obtained using the chi-square distribution.

---

### Bootstrap Confidence Interval

To quantify uncertainty without relying solely on asymptotic theory,
we also construct a parametric bootstrap confidence interval for $\psi$.
Bootstrap samples are generated according to

\[
T^* \sim \text{Binomial}(n,\hat{\pi}),
\]

and the corresponding estimator $\hat{\psi}^*$ is calculated for each
bootstrap sample. The empirical $2.5\%$ and $97.5\%$ quantiles of the
bootstrap distribution provide a $95\%$ percentile confidence interval
for $\psi$.

## Bayesian Inference
Detail the Bayesian approach.

# Results
Present your findings. 


```{r}
# Data from Pfizer trial
n <- 170        # total infections
t <- 8          # infections in vaccine group

# -------------------------
# MLE for pi
# -------------------------
pi_hat <- t/n
pi_hat

# -------------------------
# MLE for vaccine efficacy psi
# psi = (1 - 2*pi)/(1 - pi)
# -------------------------
psi_hat <- (1 - 2*pi_hat)/(1 - pi_hat)
psi_hat


# -------------------------
# Log-likelihood function
# -------------------------
loglik <- function(pi, t, n){
  if(pi <= 0 | pi >= 1) return(-Inf)
  dbinom(t, size=n, prob=pi, log=TRUE)
}

# evaluate likelihood across grid
pi_vals <- seq(0.001,0.999,length=1000)
ll_vals <- sapply(pi_vals, loglik, t=t, n=n)

plot(pi_vals, ll_vals, type="l",
     xlab="pi",
     ylab="log-likelihood",
     main="Log-Likelihood Function")


# -------------------------
# Large sample CI for pi
# -------------------------
se_pi <- sqrt(pi_hat*(1-pi_hat)/n)

ci_pi <- pi_hat + c(-1,1)*qnorm(0.975)*se_pi
ci_pi


# -------------------------
# Convert CI to psi scale
# -------------------------
psi_transform <- function(pi){
  (1-2*pi)/(1-pi)
}

ci_psi <- psi_transform(ci_pi)
ci_psi


# -------------------------
# Likelihood Ratio Test
# H0: psi = 0.3
# -------------------------

psi0 <- 0.3

# convert psi0 to pi0
pi0 <- (1 - psi0)/(2 - psi0)

# likelihood ratio statistic
W <- 2*(loglik(pi_hat,t,n) - loglik(pi0,t,n))
W

# p-value from chi-square distribution
p_value <- 1 - pchisq(W, df=1)
p_value


# -------------------------
# Parametric bootstrap CI
# -------------------------
B <- 10000

boot_pi <- rbinom(B, n, pi_hat)/n
boot_psi <- psi_transform(boot_pi)

boot_ci <- quantile(boot_psi, c(0.025,0.975))
boot_ci
```


# Discussion / Conclusion
Discuss / conclude here.

# Bibliography

Brown, B. (2024). *Lecture Title*. Lecture slides, Course Name, University Name.

Doe, J. (2020). Title of the Paper. *Journal Name*, 12(3), 45-67.

Last, F., & Last, F. (2025). *Book Title*. Publisher.

Smith, A., & Johnson, C. (2023). *Title of the Online Article*. Retrieved from https://www.example.com.

# Appendix

## Code
```{r ref.label = "important_R_code", eval=FALSE}
```

## Proofs
If applicable, include detailed mathematical derivations or additional theoretical explanations.
