---
title: "Your Paper Title Here"
author: "Your Name(s)"
date: Date
output: pdf_document
urlcolor: blue
header-includes:
- \usepackage{amsmath,amsfonts,amssymb}
- \usepackage{setspace} \doublespacing
fontsize: 11pt
---

```{r setup, include=FALSE}
library(ggplot2)
library(LearnBayes)
library(HDInterval)
```

# Abstract
Write your abstract here.

# Keywords
*Keyword 1*, *Keyword 2*, *Keyword 3*, *Keyword 4*

\newpage

# Introduction / Background
The December 2020 Emergency Use Authorization (EUA) of the BNT162b2 vaccine marked a pivotal shift in the global effort to mitigate the SARS-CoV-2 pandemic. Developed through a collaboration between Pfizer and BioNTech, this two-dose regimen represented a departure from traditional vaccine technologies. As noted in contemporary scientific literature, while viral vector approaches were utilized by manufacturers like Janssen and AstraZeneca, BNT162b2 utilized lipid nanoparticle-formulated nucleoside-modified RNA (mRNA) to induce an immune response against the viral spike protein. To evaluate the safety and clinical impact of this technology, a rigorous, multinational, placebo-controlled efficacy study was conducted and published by Polack et al. (2020). This trial followed a 1:1 randomization ratio, assigning over 34,000 participants to either the vaccine or placebo groups. The study was designed to reach statistical maturity upon the occurrence of 170 laboratory-confirmed COVID-19 cases in participants without prior evidence of infection. The resulting data demonstrated a significant disparity in outcomes. Of the 170 confirmed cases, 8 occurred in the BNT162b2 cohort ($n = 17,411$), while 162 occurred in the placebo cohort ($n = 17,511$). This results in an observed infection rate of approximately 0.046% for the vaccine group compared to 0.925% for the placebo group.
```{r}
trial_matrix <- matrix(c(8, 17411-8, 162, 17511-162), 
                       nrow = 2, 
                       dimnames = list(Outcome = c("Infected", "Healthy"),
                                       Group = c("BNT162b2", "Placebo")))

rates <- trial_matrix["Infected", ] / colSums(trial_matrix)
barplot(rates, 
        main="Infection Rate by Group", 
        ylab="Proportion of Cases", 
        col=c("skyblue", "salmon"),
        ylim=c(0, 0.01))
array <- as.array(trial_matrix)
mosaicplot(array, main="Mosaic Plot of Distribution", sub = NULL, xlab = NULL, ylab = NULL,
           sort = NULL, off = NULL, dir = NULL,
           color = c("skyblue","salmon"), shade = FALSE, margin = NULL,
           cex.axis = 0.66, las = par("las"), border = NULL)

```
This analysis seeks to evaluate the efficacy of the vaccine based on these findings. Specifically, we test whether the vaccine efficacy ($\psi$) is significantly greater than the 30% threshold required for regulatory approval. We formulate our research goal through the hypotheses $H_0: \psi = 0.3$ versus $H_1: \psi > 0.3$.




```{r label="important_R_code", eval=TRUE, echo=FALSE}
# You can reference your code in the appendix (sample here).
```

# Statistical Methods

## Model
Describe the statistical model used.

## Likelihood Inference
Detail the likelihood approach.

## Bayesian Inference

```{r, echo=FALSE, message=FALSE}

# 1. Prior Elicitation
# We want P(psi > 0.3) = 0.05. 
# Using the transformation pi = (1-psi)/(2-psi):
# If psi = 0.3, then pi = (1-0.3)/(2-0.3) = 0.7/1.7 = 0.41176
# We look for a Beta(a, b) where the 5th percentile is 0.41176 and the median is 0.5.

prior_params <- beta.select(quantile1 = list(p = 0.05, x = 0.7/1.7), 
                            quantile2 = list(p = 0.5, x = 0.5))

a_prior <- prior_params[1] # This will be 43.04
b_prior <- prior_params[2] # This will be 43.04

# 2. Observed Data from Instructions
t_obs <- 8      # Infections in vaccine group
n_total <- 170  # Total infections (8 + 162)

# 3. Posterior Calculation
# Using Conjugacy: Beta(a + t, b + n - t)
a_post <- a_prior + t_obs
b_post <- b_prior + (n_total - t_obs)

# 4. Calculating Results for the "Results" section (Calculated here for accuracy)
post_median_pi <- qbeta(0.5, a_post, b_post)
post_hdi_pi <- hdi(qbeta, credMass = 0.95, shape1 = a_post, shape2 = b_post)

# Convert pi results back to Efficacy (psi) scale
post_median_psi <- (1 - 2 * post_median_pi) / (1 - post_median_pi)
post_hdi_psi <- (1 - 2 * post_hdi_pi) / (1 - post_hdi_pi)

# 5. Hypothesis Test: P(psi > 0.3 | data)
# This is equivalent to P(pi < 0.41176 | data)
prob_efficacy_gt_30 <- pbeta(0.7/1.7, a_post, b_post)
```
The Bayesian approach requires the specification of a prior distribution for $\pi$ that represents our state of knowledge before observing the trial data. We adopt a conjugate Beta prior, $\pi \sim Beta(a, b)$, which is mathematically advantageous as it ensures the posterior distribution will also belong to the Beta family.To determine the values of the hyperparameters $a$ and $b$, we perform a prior elicitation based on a "pessimistic" perspective of vaccine performance. We assume a prior median for $\pi$ of $0.5$, reflecting an initial belief that the vaccine and placebo groups are equally likely to result in an infection. Furthermore, we reflect regulatory skepticism by setting only a 5% prior probability that the vaccine efficacy meets the 30% threshold ($P(\psi \geq 0.30) = 0.05$). Converting this efficacy threshold to the $\pi$ scale using the transformation $\pi = \frac{1-\psi}{2-\psi}$, we target a distribution where the 5th percentile is approximately $0.4118$. Solving these constraints yields the parameters:$$a = 43.04, \quad b = 43.04$$The likelihood of observing $t$ infections in the vaccine group out of $n$ total cases is given by the binomial probability mass function:$$f(t | \pi) = \binom{n}{t} \pi^t (1-\pi)^{n-t}$$Applying Bayes’ Theorem, the posterior distribution $f(\pi | t)$ is proportional to the product of the likelihood and the prior. Due to the conjugacy of the Beta-binomial model, the posterior distribution for $\pi$ is updated by adding the observed successes and failures to the prior parameters:$$\pi | t \sim Beta(a + t, b + n - t)$$With our observed data of $t=8$ and $n=170$, the posterior becomes $\pi | t \sim Beta(51.04, 205.04)$. Following the calculation of this distribution, we transform the results back to the $\psi$ scale. This allows for the determination of the posterior median for vaccine efficacy and a 95% Highest Posterior Density Interval (HPDI) to quantify uncertainty. Finally, the posterior probability $P(\psi > 0.30 | t)$ is calculated to evaluate the strength of evidence in favor of the research hypothesis.
```{r}

# Define parameters
a_prior <- 43.04
b_prior <- 43.04
t_obs <- 8
n_total <- 170
a_post <- a_prior + t_obs
b_post <- b_prior + (n_total - t_obs)

# Generate data for plotting
x <- seq(0, 1, length.out = 500)
df <- data.frame(
  pi = rep(x, 2),
  pdf = c(dbeta(x, a_prior, b_prior), dbeta(x, a_post, b_post)),
  distribution = rep(c("Prior", "Posterior"), each = 500)
)

# Plot
ggplot(df, aes(x = pi, y = pdf, color = distribution)) +
  geom_line(linewidth = 1) +
  labs(title = "Prior and Posterior Distribution of pi",
       x = expression(pi),
       y = "Density") +
  theme_minimal() +
  scale_color_manual(values = c("Prior" = "gray60", "Posterior" = "firebrick"))
```


# Results
The Bayesian analysis provides a probabilistic assessment of the vaccine efficacy parameter, $\psi$, derived from the posterior distribution of $\pi$. Using the pessimistic prior distribution ($Beta(43.04, 43.04)$) and the trial data of 8 vaccine infections out of 170 total cases, the posterior distribution for $\pi$ was found to be $Beta(51.04, 205.04)$.The posterior median for the probability that an infected individual belonged to the vaccine group, $\pi$, is 0.199. When transformed to the vaccine efficacy scale using the relationship $\psi = \frac{1-2\pi}{1-\pi}$, this yields a posterior median vaccine efficacy of 75.1%. The uncertainty in this estimate is characterized by a 95% Highest Posterior Density Interval (HPDI) for $\psi$ of [66.5%, 82.2%].To evaluate the research hypothesis $H_1: \psi > 0.3$, the posterior probability was calculated. The analysis indicates that the posterior probability of the vaccine efficacy exceeding the 30% regulatory threshold, $P(\psi > 0.30 | \text{data})$, is greater than 0.9999. These inferential statistics are summarized in Table 2.

Table 2: Bayesian Posterior Summaries for Vaccine Efficacy ($\psi$)| Metric | Posterior Estimate || :--- | :--- || Posterior Median ($\psi$) | 75.1% || 95% HPDI ($\psi$) | [66.5%, 82.2%] || $P(\psi > 0.30 \mid \text{data})$ | > 0.9999 |

# Discussion / Conclusion
Discuss / conclude here.

# Bibliography

Grove, R. (2026). *Physicians-Health-Study.pdf*. Lecture slides, STAT 342, University of Washington.

Grove, R. (2026). *bayes-inference.pdf*. Lecture slides, STAT 342, University of Washington.

Grove, R. (2026). *likelihood-inference.pdf*. Lecture slides, STAT 342, University of Washington.

Polack, F. P., Thomas, S. J., Kitchin, N., Absalon, J., Gurtman, A., Lockhart, S., . . . Gruber, W. C. (2020). Safety and Efficacy of the BNT162b2 mRNA Covid-19 Vaccine. New England Journal of Medicine, 383(27), 2603-2615.

# Appendix

## Code
```{r ref.label = "important_R_code", eval=FALSE}
```

## Proofs
If applicable, include detailed mathematical derivations or additional theoretical explanations.
